---
title: "01-Simple models"
author: "Francisco E. Fonturbel"
date: "19/nov/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## Anthropogenic effects on chungungo's abundance

Here we will explore some anthropogenic variables as potential drivers os abundance change on this species

```{r data}
#loading libraries
library(mgcv)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(MuMIn)
library(MCMCglmm)
library(ggplot2)
source("multiplot.R")

#loading data
data<-read.table("data/data.txt", header = T)
attach(data)
```

Let's do some nice ggplots to examine our data before fitting fancy models:

```{r plots_oc, echo=FALSE, fig.height=12, fig.width=10, message=FALSE, warning=FALSE}
a4<-ggplot(data=data, aes(x=p_ratio, y=PLAND)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "darkblue") +
  labs(x="D. gliroides occurrence", y="Forest %") +
  ggtitle("PLAND") +
  theme_classic() +
  My_Theme
a6<-ggplot(data=data, aes(x=p_ratio, y=NP)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "blueviolet") +
  labs(x="D. gliroides occurrence", y="Number of patches") +
  ggtitle("NP") +
  theme_classic() +
  My_Theme
b4<-ggplot(data=data, aes(x=p_ratio, y=contig_wg)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "firebrick1") +
  labs(x="D. gliroides occurrence", y="Contiguity Index (wg mean)") +
  ggtitle("CONTIG (wg mean)") +
  theme_classic() +
  My_Theme
c8<-ggplot(data=data, aes(x=p_ratio, y=ENN_wg)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "darkorange1") +
  labs(x="D. gliroides occurrence", y="ENN (wg mean)") +
  ggtitle("ENN (wg mean)") +
  theme_classic() +
  My_Theme
d2<-ggplot(data=data, aes(x=p_ratio, y=CONNECT_250m)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "chartreuse3") +
  labs(x="D. gliroides occurrence", y="Connectance Index") +
  ggtitle("CONNECT 250m") +
  theme_classic() +
  My_Theme

multiplot(a4, b4, d2, a6, c8, cols=2)
```

## Models

To do this, we will use a Poisson GLMM, using site and sampling month as random factors.

```{r model1}
mod1<-glmer(abund~people+boats+pets+algae+(1|site)+(1|month), family = "poisson", na.action = "na.fail")
summary(mod1)
```

Well, it seems that none of our predictor variables are significant... Also, R warns us that we have _singular fit_ (this [article](https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#singular-models-random-effect-variances-estimated-as-zero-or-correlations-estimated-as---1) to solve that)

Let's try some multi-model inference to examine the relative support of each variable to our dataset

```{r mumin1}
mm1<-dredge(mod1, rank = BIC)
subset(mm1, delta <5)
get.models(mm1, subset = delta < 2)
imp.mm1<-importance(mm1)
barplot(t(imp.mm1), main="Predictors", col="black", xlab="Variable", ylab="Relative support")
```

Well, it seems that all predictor variables have similar importance.

Following Ben Bolker's advice, we can avoid singular fit by removing those random effects with a very litle contribution (in this case, _month_). Let's try fitting a new model using only _site_ as random effect:

```{r model2}
mod2<-glmer(abund~people+boats+pets+algae+(1|site), family = "poisson", na.action = "na.fail")
summary(mod2)
```

Well, there is no singular fit now, BUT none of our predictor variables is significant (doh!)

Let's examine multi-model inference for this new model as well:

```{r mumin2}
mm2<-dredge(mod2, rank = BIC)
subset(mm2, delta <5)
get.models(mm2, subset = delta < 2)
imp.mm2<-importance(mm2)
barplot(t(imp.mm2), main="Predictors", col="black", xlab="Variable", ylab="Relative support")
```

Pretty much the same than above.

### A Bayesian approach

As we have a relatively small (N = 34) sample size, we'll try now a Bayesian approach (based on Dimas' code) to test the same but under a Bayesian framework based on Monte Carlo Markov Chains

```{r mcmc1}
mod3<-MCMCglmm(abund~people+boats+pets+algae, random = ~ site, data=data, verbose = FALSE)
summary(mod3)
```

OK, it seems that there is no effect at all, as both approaches showed consistent results.

## Session

```{r session, echo=T}
sessionInfo()
```