---
title: "01-Simple models - disturbance index"
author: "Francisco E. Fonturbel"
date: "09/mar/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## Anthropogenic effects on chungungo's abundance

Here we will explore some anthropogenic variables as potential drivers of abundance change on this species

```{r data, include=FALSE}
#loading libraries
library(mgcv)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(MuMIn)
library(MCMCglmm)
library(ggplot2)
source("multiplot.R")

#loading data
data<-read.table("data/data4.txt", header = T)
attach(data)

#ggplot grooming
My_Theme = theme(
  axis.title.x = element_text(size = 18,face="bold"),
  axis.text.x = element_text(size = 16),
  axis.text.y = element_text(size = 16),
  axis.title.y = element_text(size = 18,face="bold"),
  legend.text = element_text(size = 14),
  legend.title = element_text(size = 14,face="bold"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(size = 24,face = "bold", hjust=0.5))
```

Let's do some nice ggplots to examine our data before fitting fancy models:

```{r plots_oc, echo=FALSE, fig.height=12, fig.width=10, message=FALSE, warning=FALSE}
a1<-ggplot(data=data, aes(x=distindx, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "dodgerblue2") +
  labs(x="Dist index", y="Abundance") +
  ggtitle("Disturbance index") +
  theme_classic() +
  My_Theme

multiplot(a1, cols=1)
```

There is a trend but the error is HUGE!

## Models

To do this, we will use a simple GLMM, using site and sampling month as random factors.

```{r model1}
mod1<-lmer(abund~distindx+(1|site)+(1|month), na.action = "na.fail")
summary(mod1, ddf="Kenward-Roger")
```

Well, it seems that the new index is significant...

Following Ben Bolker's advice, we can avoid singular fit by removing those random effects with a very litle contribution (in this case, _month_). Let's try fitting a new model using only _site_ as random effect:

```{r model2}
mod2<-lmer(abund~distindx+(1|site), na.action = "na.fail")
summary(mod2, ddf="Kenward-Roger")
```

Well, there is no singular fit now, BUT the disturbance index is not significant as well (doh!)


```{r models}
mods<-glm(log(abund+1)~log(distindx+1))
summary(mods)
```


### A Bayesian approach

As we have a relatively small (N = 34) sample size, we'll try now a Bayesian approach (based on Dimas' code) to test the same but under a Bayesian framework based on Monte Carlo Markov Chains

```{r mcmc1}
mod3<-MCMCglmm(abund~distindx, random = ~ site, data=data, verbose = FALSE)
summary(mod3)
```

OK, it seems that there is no effect at all, as all approaches showed consistent results.

## Session

```{r session, echo=T}
sessionInfo()
```