---
title: "02-Habitat"
author: "Francisco E. Fonturbel"
date: "09/mar/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## Anthropogenic effects on chungungo's abundance

Here we will explore some anthropogenic variables as potential drivers os abundance change on this species

```{r data, include=FALSE}
#loading libraries
library(ggplot2)
library(ggsci)
library(vegan)
source("multiplot.R")

#loading data
data<-read.table("data/data.txt", header = T)
attach(data)

hab<-read.table("data/habitat.txt", header = T)

#ggplot grooming
My_Theme = theme(
  axis.title.x = element_text(size = 18,face="bold"),
  axis.text.x = element_text(size = 16),
  axis.text.y = element_text(size = 16),
  axis.title.y = element_text(size = 18,face="bold"),
  legend.text = element_text(size = 14),
  legend.title = element_text(size = 14,face="bold"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(size = 24,face = "bold", hjust=0.5))
```

We have characterized six sites where _Lontra felina_ is present.

First, are those sites different in terms of habitat features? Let's do an `ANOSIM` and a `nMDS`  for that:

### ANOSIM

```{r anosim, eval=TRUE, echo=FALSE, fig.width=8}
habitat<-hab[,3:10]
habdist<-vegdist(habitat)
  
hab.anosim<-anosim(habdist, hab$site, permutations = 999, distance = "bray")
summary(hab.anosim)
plot(hab.anosim, xlab="Site", ylab="Rank distance")
```

No, the six sites are similar in terms of habitat.

### nMDS plot

```{r nmds}
hab.mds<-metaMDS(habdist, distance = "bray")

hab.scores<-as.data.frame(scores(hab.mds))
hab.scores$site<-rownames(hab.scores)
hab.scores$site<-hab$site

grp.G1 <- hab.scores[hab.scores$site == "S1", ][chull(hab.scores[hab.scores$site == 
    "S1", c("NMDS1", "NMDS2")]), ]  # hull values for site 1
grp.G2 <- hab.scores[hab.scores$site == "S2", ][chull(hab.scores[hab.scores$site == 
    "S2", c("NMDS1", "NMDS2")]), ]  # hull values for site 2
grp.G3 <- hab.scores[hab.scores$site == "S3", ][chull(hab.scores[hab.scores$site == 
    "S3", c("NMDS1", "NMDS2")]), ]  # hull values for site 3
grp.G4 <- hab.scores[hab.scores$site == "S4", ][chull(hab.scores[hab.scores$site == 
    "S4", c("NMDS1", "NMDS2")]), ]  # hull values for site 4
grp.G5 <- hab.scores[hab.scores$site == "S5", ][chull(hab.scores[hab.scores$site == 
    "S5", c("NMDS1", "NMDS2")]), ]  # hull values for site 5
grp.G6 <- hab.scores[hab.scores$site == "S6", ][chull(hab.scores[hab.scores$site == 
    "S6", c("NMDS1", "NMDS2")]), ]  # hull values for site 6


hull.hab <- rbind(grp.G1, grp.G2, grp.G3, grp.G4, grp.G5, grp.G6)  #combine groups

hab.nmds.plot<-ggplot() + 
  geom_polygon(data=hull.hab,aes(x=NMDS1,y=NMDS2,fill=site,group=site),alpha=0.30) + # add the convex hulls
  geom_point(data=hab.scores,aes(x=NMDS1,y=NMDS2,shape=site,colour=site),size=4) + # add the point markers
  coord_equal() +
  scale_fill_startrek() +
  scale_color_startrek() +
  theme_bw() + 
  labs(x="nMDS 1", y="nMDS 2") +
  theme(axis.text.x = element_text(size = 16),  # remove x-axis text
        axis.text.y = element_text(size = 16), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 16))
hab.nmds.plot
#Export in landscape mode 7x10

#Obtaining nMDS stress
hab.mds$stress
```

Well, they are pretty much the same. Stress values are quite good!

### MANOVA

In a last attempt to test for habitat differences, we will perform a multivariate analysis of variance using R's `manova` function.

```{r manowar}
hab_diff<-manova(as.matrix(habitat)~hab$site)
summary(hab_diff)
summary.aov(hab_diff)
```

Only distance showed a significant difference.

## Session

```{r session, echo=T}
sessionInfo()
```