---
title: "02-Habitat (2024 version)"
author: "Francisco E. Fonturbel, Guido Pavez"
date: "01/Feb/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## Variation in otter habitat variables.


```{r data, include=FALSE}
#loading libraries
library(ggplot2)
library(ggsci)
library(vegan)
source("multiplot.R")

#loading data
data<-read.table("data/Habitat_2024.txt", header = T)
attach(data)

#ggplot grooming
My_Theme = theme(
  axis.title.x = element_text(size = 18,face="bold"),
  axis.text.x = element_text(size = 16),
  axis.text.y = element_text(size = 16),
  axis.title.y = element_text(size = 18,face="bold"),
  legend.text = element_text(size = 14),
  legend.title = element_text(size = 14,face="bold"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(size = 24,face = "bold", hjust=0.5))
```

We have characterized six sites where _Lontra felina_ is present. Here we have updated data based on drone estimations (more accurate than visual estimations that we had).

First, are those sites different in terms of habitat features? Let's do an `ANOSIM` and a `nMDS`  for that:

### ANOSIM

```{r anosim, eval=TRUE, echo=FALSE, fig.width=8}
habitat<-data[,3:8]

habdist<-vegdist(habitat)
  
hab.anosim<-anosim(habdist, data$site, permutations = 999, distance = "jaccard")
summary(hab.anosim)
plot(hab.anosim, xlab="Site", ylab="Rank distance")
```

It seems that Site 6 is different than the others in terms of habitat structure.

### nMDS plot

```{r nmds}
hab.mds<-metaMDS(habdist, distance = "bray")

hab.scores<-as.data.frame(scores(hab.mds))
hab.scores$site<-rownames(hab.scores)
hab.scores$site<-data$site

grp.G1 <- hab.scores[hab.scores$site == "S1", ][chull(hab.scores[hab.scores$site == 
    "S1", c("NMDS1", "NMDS2")]), ]  # hull values for site 1
grp.G2 <- hab.scores[hab.scores$site == "S2", ][chull(hab.scores[hab.scores$site == 
    "S2", c("NMDS1", "NMDS2")]), ]  # hull values for site 2
grp.G3 <- hab.scores[hab.scores$site == "S3", ][chull(hab.scores[hab.scores$site == 
    "S3", c("NMDS1", "NMDS2")]), ]  # hull values for site 3
grp.G4 <- hab.scores[hab.scores$site == "S4", ][chull(hab.scores[hab.scores$site == 
    "S4", c("NMDS1", "NMDS2")]), ]  # hull values for site 4
grp.G5 <- hab.scores[hab.scores$site == "S5", ][chull(hab.scores[hab.scores$site == 
    "S5", c("NMDS1", "NMDS2")]), ]  # hull values for site 5
grp.G6 <- hab.scores[hab.scores$site == "S6", ][chull(hab.scores[hab.scores$site == 
    "S6", c("NMDS1", "NMDS2")]), ]  # hull values for site 6


hull.hab <- rbind(grp.G1, grp.G2, grp.G3, grp.G4, grp.G5, grp.G6)  #combine groups

hab.nmds.plot<-ggplot() + 
  geom_polygon(data=hull.hab,aes(x=NMDS1,y=NMDS2,fill=site,group=site),alpha=0.30) + # add the convex hulls
  geom_point(data=hab.scores,aes(x=NMDS1,y=NMDS2,shape=site,colour=site),size=4) + # add the point markers
  coord_equal() +
  scale_fill_startrek() +
  scale_color_startrek() +
  theme_bw() + 
  labs(x="nMDS 1", y="nMDS 2") +
  theme(axis.text.x = element_text(size = 16),  # remove x-axis text
        axis.text.y = element_text(size = 16), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 16))
hab.nmds.plot

#Obtaining nMDS stress
hab.mds$stress
```

Well, they are pretty much the same. Stress values are quite good!

### MANOVA

In a last attempt to test for habitat differences, we will perform a multivariate analysis of variance using R's `manova` function.

```{r manowar}
hab_diff<-manova(as.matrix(habitat)~data$site)
summary(hab_diff)
summary.aov(hab_diff)
```
OK, habitat structure is indeed different among sites. The variables explaining those differences are: smallr, larger, reef, and slope.

## Session

```{r session, echo=T}
sessionInfo()
```