---
title: "02-GLMM 2024"
author: "Francisco E. Fonturbel"
date: "31/ene/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## Anthropogenic effects on chungungo's abundance

Here we will explore some anthropogenic variables as potential drivers of abundance change on this species

```{r data, include=FALSE}
#loading libraries
library(mgcv)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(MuMIn)
library(MCMCglmm)
library(ggplot2)
library(AER)
source("multiplot.R")

#loading data
data<-read.table("data/Data_2024.txt", header = T)
attach(data)


#ggplot grooming
My_Theme = theme(
  axis.title.x = element_text(size = 10),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  legend.text = element_text(size = 12),
  legend.title = element_text(size = 12),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(size = 10,face = "bold", hjust=0.5))
```

Let's do some nice ggplots to examine our data before fitting fancy models:

```{r plots_oc, echo=FALSE, fig.height=12, fig.width=10, message=FALSE, warning=FALSE}
a1<-ggplot(data=data, aes(x=people, y=abund)) +
  geom_smooth(linewidth=1, span = 2, method = "glm", color = "black") +
  labs(x="Number of people", y="Relative abundance (ind/km)") +
  theme_classic() +
  My_Theme
  
a2<-ggplot(data=data, aes(x=boats, y=abund)) +
  geom_smooth(linewidth=1, span = 2, method = "glm", color = "black") +
  labs(x="Number of boats", y="Relative abundance (ind/km)") +
  theme_classic() +
  My_Theme
  
a3<-ggplot(data=data, aes(x=pets, y=abund)) +
  geom_smooth(linewidth=1, span = 2, method = "glm", color = "black") +
  labs(x="Number of pets", y="Relative abundance (ind/km)") +
  theme_classic() +
  My_Theme
a4<-ggplot(data=data, aes(x=index_1, y=abund)) +
  geom_smooth(linewidth=1, span = 2, method = "glm", color = "black") +
  labs(x="Human Disturbance Index", y="Relative abundance (ind/km)") +
  theme_classic() +
  My_Theme

multiplot(a1,a2,a3,a4, cols=2)
```

##How is our **response varible** distributed?

```{r hist, echo=FALSE}
hist(abund, col = "slateblue2", main = "Response variable", xlab = "Relative abundance", ylab = "Frequency")
```

##Well, our response variable is not normal, as abundance is a count variable with positive integers. Therefore, we should use _Poisson_ or _negative binomial_ distributions instead.

## Models

#To do this, we will use a Poisson GLMM, using sampling site as random factors.

```{r model1}
mod1<-glmer(abund~people+boats+pets+(1|site), na.action = "na.fail", family = "poisson")
summary(mod1)
```

##Well, the variable number of people is significant

#GLMM para el índice de perturbación antrópica

```{r model1}

mod2<-glmer(abund~index_1+(month|year), na.action = "na.fail", family = "poisson")
summary(mod2)

```

#Singular fit too :C

Let's try some multi-model inference to examine the relative support of each variable to our dataset


```{r mumin1}
mm1<-dredge(mod1, rank = BIC)
subset(mm1, delta <5)
get.models(mm1, subset = delta < 2)
imp.mm1<-sw(mm1)
barplot(t(imp.mm1), main="Predictors", col="black", xlab="Variable", ylab="Relative support")
```

Following Ben Bolker's advice, we can avoid singular fit by removing those random effects with a very litle contribution (in this case, _month_). Let's try fitting a new model using only _site_ as random effect:

```{r model2}
mod3<-glmer(abund~people+boats+pets+(1 | site), na.action = "na.fail", family = "poisson")
summary(mod3)
```

Probé dejando solo el Mes como factor aleatorio (sigue arrojando Singular fit) y solo el Sitio como factor aleatorio (ya no arroja Singular fit pero pierde significancia)

Let's examine multi-model inference for this new model as well:

```{r mumin2}
mm2<-dredge(mod2, rank = BIC)
subset(mm2, delta <5)
get.models(mm2, subset = delta < 2)
imp.mm2<-importance(mm2)
barplot(t(imp.mm2), main="Predictors", col="black", xlab="Variable", ylab="Relative support")
```

Los siguientes modelos quedaron del Script original:

Pretty much the same than above.

### A Bayesian approach

We'll try now a Bayesian approach (based on Dimas' code) to test the same but under a Bayesian framework based on Monte Carlo Markov Chains

```{r mcmc1}
mod4<-MCMCglmm(abund~people+boats+pets, random = ~ site, data=data, verbose = FALSE)
summary(mod4)
```

Al correr esto da casi significativo la variable Personas.

## Session

```{r session, echo=T}
sessionInfo()
```