---
title: "03-Models for 2yr dataset"
author: "Francisco E. Fonturbel"
date: "28/jun/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## Anthropogenic effects on chungungo's abundance

Here we will explore some anthropogenic variables as potential drivers of abundance change on this species

```{r data, include=FALSE}
#loading libraries
library(mgcv)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(MuMIn)
library(MCMCglmm)
library(ggplot2)
source("multiplot.R")

#loading data
data<-read.table("data/data_2023.txt", header = T)
attach(data)

#ggplot grooming
My_Theme = theme(
  axis.title.x = element_text(size = 18,face="bold"),
  axis.text.x = element_text(size = 16),
  axis.text.y = element_text(size = 16),
  axis.title.y = element_text(size = 18,face="bold"),
  legend.text = element_text(size = 14),
  legend.title = element_text(size = 14,face="bold"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(size = 24,face = "bold", hjust=0.5))
```

Let's do some nice ggplots to examine our data before fitting fancy models:

```{r plots_oc, echo=FALSE, fig.height=12, fig.width=6, message=FALSE, warning=FALSE}
a1<-ggplot(data=data, aes(x=people, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "darkblue") +
  labs(x="People", y="Abundance") +
  ggtitle("People") +
  theme_classic() +
  My_Theme
a2<-ggplot(data=data, aes(x=boats, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "blueviolet") +
  labs(x="Boats", y="Abundance") +
  ggtitle("Boats") +
  theme_classic() +
  My_Theme
a3<-ggplot(data=data, aes(x=pets, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "firebrick1") +
  labs(x="Pets", y="Abundance") +
  ggtitle("Pets") +
  theme_classic() +
  My_Theme

multiplot(a1, a2, a3, cols=1)
```


## Individual models

Although is not the best option, we will try to fit some glms for each individual predictor variable depicted in the plots above.

### People
```{r modelp}
modp<-glm(abund~people)
summary(modp)
```

### Boats
```{r modelb}
modb<-glm(abund~boats)
summary(modb)
```

### Pets
```{r modelt}
modt<-glm(abund~pets)
summary(modt)
```


## Models

To do this, we will use a Poisson GLMM, using site and sampling month as random factors.

```{r model1}
mod1<-lmer(abund~people+boats+pets+(1|site)+(month|year), na.action = "na.fail")
summary(mod1)
```

Well, it seems that none of our predictor variables are significant... 

```{r mumin1}
mm1<-dredge(mod1, rank = BIC)
subset(mm1, delta <2)
get.models(mm1, subset = delta < 2)
imp.mm1<-sw(mm1)
barplot(t(imp.mm1), main="Predictors", col="black", xlab="Variable", ylab="Relative support")
```

Well, we have a trend here! but relative support is way too low.


### A Bayesian approach

As we have a relatively small (N = 34) sample size, we'll try now a Bayesian approach (based on Dimas' code) to test the same but under a Bayesian framework based on Monte Carlo Markov Chains

```{r mcmc1}
mod2<-MCMCglmm(abund~people+boats+pets, random = ~ site, data=data, verbose = FALSE)
summary(mod2)
```

OK, it seems that there is no effect at all, as both approaches showed consistent results.


## Using disturbance indices

We made two disturbance indices upon the disturbance levels measured. Now, we will use those indices instead of the raw variables... let's see:

#### Index 1

```{r mind1}
mod3<-lmer(abund~index_1+(1|site)+(month|year), na.action = "na.fail")
summary(mod3)
```

#### Index 2

```{r mind2}
mod4<-lmer(abund~index_2+(1|site)+(month|year), na.action = "na.fail")
summary(mod4)
```


## Session

```{r session, echo=T}
sessionInfo()
```