---
title: "03-Models for 2yr dataset"
author: "Francisco E. Fonturbel"
date: "16/ago/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

## Anthropogenic effects on chungungo's abundance

Here we will explore some anthropogenic variables as potential drivers of abundance change on this species

```{r data, include=FALSE}
#loading libraries
library(mgcv)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(MuMIn)
library(MCMCglmm)
library(ggplot2)
library(AER)
source("multiplot.R")

#loading data
data<-read.table("data/data_2023.txt", header = T)
attach(data)

#ggplot grooming
My_Theme = theme(
  axis.title.x = element_text(size = 18,face="bold"),
  axis.text.x = element_text(size = 16),
  axis.text.y = element_text(size = 16),
  axis.title.y = element_text(size = 18,face="bold"),
  legend.text = element_text(size = 14),
  legend.title = element_text(size = 14,face="bold"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(size = 24,face = "bold", hjust=0.5))
```

Let's do some nice ggplots to examine our data before fitting fancy models:

```{r plots_oc, echo=FALSE, fig.height=12, fig.width=6, message=FALSE, warning=FALSE}
a1<-ggplot(data=data, aes(x=people, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "darkblue") +
  labs(x="People", y="Abundance") +
  ggtitle("People") +
  theme_classic() +
  My_Theme
a2<-ggplot(data=data, aes(x=boats, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "blueviolet") +
  labs(x="Boats", y="Abundance") +
  ggtitle("Boats") +
  theme_classic() +
  My_Theme
a3<-ggplot(data=data, aes(x=pets, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "firebrick1") +
  labs(x="Pets", y="Abundance") +
  ggtitle("Pets") +
  theme_classic() +
  My_Theme

multiplot(a1, a2, a3, cols=1)
```


How is our **response varible** distributed?

```{r hist, echo=FALSE}
hist(abund, col = "slateblue2", main = "Response variable", xlab = "Abundance", ylab = "Frequency")
```

Well, our response variable is not normal, as abundance is a count variable with positive integers. Therefore, we should use _Poisson_ or _negative binomial_ distributions instead.


## Individual models

Although is not the best option, we will try to fit some glms for each individual predictor variable depicted in the plots above.

### People
```{r modelp}
modp<-glm(abund~people, family = "poisson")
summary(modp)
```

Let's test our count variable for overdispersion before continuing...

```{r disper}
dispersiontest(modp)
```

No overdispersion, we can go with Poisson distribution hereafter.

### Boats
```{r modelb}
modb<-glm(abund~boats, family = "poisson")
summary(modb)
```

### Pets
```{r modelt}
modt<-glm(abund~pets, family = "poisson")
summary(modt)
```


## Models

To do this, we will use a Poisson GLMM, using sampling month as a random factor (month nested within year).

```{r model1}
mod1<-glmer(abund~people+boats+pets+(month|year), na.action = "na.fail", family = "poisson")
summary(mod1)
```

Well, it seems that none of our predictor variables are significant... 

```{r mumin1}
mm1<-dredge(mod1, rank = BIC)
subset(mm1, delta <2)
get.models(mm1, subset = delta < 2)
imp.mm1<-sw(mm1)
barplot(t(imp.mm1), main="Predictors", col="black", xlab="Variable", ylab="Relative support")
```

As predicted, people is the most relevant variable in our models.

## Using disturbance indices

We made two disturbance indices upon the disturbance levels measured.

```{r plots_ind, echo=FALSE, fig.height=12, fig.width=6, message=FALSE, warning=FALSE}
b1<-ggplot(data=data, aes(x=index_1, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "plum") +
  labs(x="Index", y="Abundance") +
  ggtitle("Index 1") +
  theme_classic() +
  My_Theme
b2<-ggplot(data=data, aes(x=index_2, y=abund)) +
  geom_smooth(size=1, span = 2, method = "glm", color = "yellowgreen") +
  labs(x="Index", y="Abundance") +
  ggtitle("Index 2") +
  theme_classic() +
  My_Theme

multiplot(b1, b2, cols=1)
```


Now, we will use those indices instead of the raw variables... let's see:

#### Index 1

```{r mind1}
mod3<-glmer(abund~index_1+(month|year), na.action = "na.fail", family = "poisson")
summary(mod3)
```

#### Index 2

```{r mind2}
mod4<-glmer(abund~index_2+(month|year), na.action = "na.fail", family = "poisson")
summary(mod4)
```

Again, nothing significant.

What if we remove the random effects?


#### Index 1 (w/o random effects)

```{r mind1nr}
mod3nr<-glm(abund~index_1, family = "poisson")
summary(mod3nr)
```

#### Index 2 (w/o random effects)

```{r mind2nr}
mod4nr<-glm(abund~index_2, family = "poisson")
summary(mod4nr)
```

Taking off the random effects yield significant results.

What if we also get rid of the random effects in the global model?

```{r model1nr}
mod1nr<-glm(abund~people+boats+pets, na.action = "na.fail", family = "poisson")
summary(mod1nr)
```

```{r mumin1nr}
mm1nr<-dredge(mod1nr, rank = BIC)
subset(mm1nr, delta <2)
get.models(mm1nr, subset = delta < 2)
imp.mm1nr<-sw(mm1nr)
barplot(t(imp.mm1nr), main="Predictors", col="violet", xlab="Variable", ylab="Relative support")
```
Now it makes more sense!

## Session

```{r session, echo=T}
sessionInfo()
```